---
title: "Intercomparison of flux-estimation approaches"
author: "Felician Heim"
format: html
editor: visual
bibliography: references.bib
---

# Preparations and load data

First, packages are loaded.

```{r}
#| echo: false
library(tidyverse)
```

Second, the different data sets are loaded, which have been preprocessed (setting date and time, converting units and renaming of objects) in `preprocessing/...`:

-   EC data output from EddyPro (3 different heights, 30m, 70m, 148m) for the sensible heat flux and $K_H$, until now only year 2020 and 30m measurement height available

-   Profile measurements (slow) from 14 levels for temperature and humidity (and potentially other gases - Picarro data)

-   L2 ICOS Ecosystem station data (EC & radiation) for evaluation of fluxes (and provides ground pressure for obtaining water vapor pressure $e$)

```{r}
#| warning: false

#load EC data output first, here only 30m until now!!!
load("../data/processed/sonic_profile_data.RData")

# load the 14 level profile data for Ta and Humidity
load("../data/processed/slow_profile_data.RData")

# Ecosystem data 30m
load("../data/processed/Eco_data_30m.RData")

```

# Bowen ratio energy balance approach

The Bowen Ratio energy balance approach does not make any assumptions about turbulence, however, makes use of data on net radiation components as well as gradient measurements from at least two heights. The calculation is based on the Bowen Ration $Bo$, being defined as

$$
Bo = \gamma\frac{\Delta T}{\Delta e} 
$$

with $\Delta T$ and $\Delta e$ denoting the air temperature and water vapor difference between two heights and The psychrometric constant $$\gamma = 0.667 hPa/K$$ for air pressure $$p = 1013.25 hPa$$ and air temperature of 20ºC (other values are given in tables, see @micromet2026, page 120/121 ).

Then, the sensible and latent heat flux can be obtained from the following equation @foken2024, page 164:

$$
Q_H = (-Q_{Net}-Q_G)*\frac{Bo}{1+Bo}
$$

$$
Q_E = \frac{-Q_{Net}-Q_G}{1+Bo}
$$

here, $Q_{Net}$, $Q_H$, $$Q_E$$, and $Q_G$ denote the net radiation, sensible heat flux, latent heat flux and ground heat flux all in $$W/m^2$$, respectively. This means, the following data has to be available to calculate the the sensible and latent heat flux (in brackets where the data is coming from):

-   net radiation (Wm2) - from the `Eco_data_30m` dataset, so the ICOS Ecosystem station

-   ground heat flux (Wm2) - from the `Eco_data_30m` dataset, so the ICOS Ecosystem station

-   The $\Delta T$ and $$\Delta e$$ from two measurement heights - from the `slow_profile_data` data set, using first 148m - 30m

    -   in order to obtain $e$ in $hPa$, one needs to convert the H2O mixing ratio in $$mmol/mol$$ first in $$kg/kg$$ by using $$m[kg/kg] = m[mmol/mol]/(1000*0.622)$$ (source/citation missing)

    -   and then, using the following equation @foken2024, page 41, Supplement 2.2, one can obtain the water vapor pressure $$e [hPa]= (m[kg/kg]*p[hPa])/(0.622)$$

    -   The gradient can then be determined as $$\Delta e = e_{148m} - e_{30m}$$

    -   This means, I need the ground pressure data from the Ecosystem Station in `Eco_data_30m` to calculate $$e$$

First, all the needed data column are merged in one data frame:

```{r}
# from the sonic data, select only H value
sonic_data_30m = sonic_profile_data%>%
  select(datetime,H_Wm2_sonic_30m)

slow_profile_data = slow_profile_data%>% select(datetime, 
                            Ta_dgC_148m, 
                            Ta_dgC_30m,
                            H2O_mmol_mol_148m, 
                            H2O_mmol_mol_30m)

Eco_data_30m = Eco_data_30m%>% select(datetime, 
                       H_Wm2, 
                       G_Wm2, 
                       LE_Wm2, 
                       R_Net_Wm2, 
                       P_ground_hPa)

  
# now merge, based on datetime
# use Eco_data as df to join in, since it contains complete 30min interval of 2020
BREB_data = Eco_data_30m%>%
  left_join(y = slow_profile_data, by = "datetime")%>%
  left_join(y = sonic_data_30m, by = "datetime")
```

Now, I can use the data to calculate the gradient $$\Delta T$$ and $$\Delta e$$, starting with the latter by make the unit conversion for mixing ratio in $$mmol/mol$$ to $$kg/kg$$, then $$e[hPa]$$.

```{r}
# calculate delta e
BREB_data = BREB_data%>%
  mutate(
    
    # convert from mmol/mol to kg/kg
    H2O_kg_kg_30m = H2O_mmol_mol_30m / 1000 * 0.622,
    H2O_kg_kg_148m = H2O_mmol_mol_148m / 1000 * 0.622,
    
    # calculate e in hPa
    e_148m_hPa = (P_ground_hPa * H2O_kg_kg_148m) / 0.622, 
    e_30m_hPa  = (P_ground_hPa * H2O_kg_kg_30m) / 0.622, 
    delta_e_hPa = e_148m_hPa - e_30m_hPa, 
    
    # calculate delta T
    delta_Ta_dgC = Ta_dgC_148m - Ta_dgC_30m
  )%>%
  # remove kg/kg columns
  select(-c(H2O_kg_kg_30m, 
            H2O_kg_kg_148m))

```

After this step, the formula for the Bowen ratio for the sensible and the latent heat can be used:

```{r}
# do the calculation for the latent and sensible heat flux
BREB_data = BREB_data%>%
  mutate(
    Bowen_ratio = 0.667 * (delta_Ta_dgC/delta_e_hPa),
    H_Wm2_BREB = (R_Net_Wm2-G_Wm2)*((Bowen_ratio)/(1+Bowen_ratio)),
    LE_Wm2_BREB = (R_Net_Wm2 - G_Wm2)/(1+Bowen_ratio)
  ) 
```

However, following @foken2024 on page 165, data must be filtered for quality reasons:

```{r}
# do quality filtering, based on Foke, Micrometeorology, page 65
BREB_data = BREB_data %>%
  mutate(
    H_Wm2_BREB  = if_else(Bowen_ratio > -1.25 
                                      & Bowen_ratio < -0.75, 
                                      true = NA, 
                                      false = H_Wm2_BREB),
    LE_Wm2_BREB = if_else(Bowen_ratio > -1.25 
                                     & Bowen_ratio < -0.75, 
                                   true = NA, 
                                   false = LE_Wm2_BREB)
  )
```

Now, the sensible and latent heat calculated using the BREB approach can be compared to the data from the Ecosystem station (EC system, processed using the ICOS protocol, 30m) and the sensible heat flux from the sonic from 30m processed by myself (EddyPro express mode for now).

Note, that this is only the data as it comes from Eddy pro (for the eddy data) - I did not do any filtering based on quality flags etc!

Using the Ecosystem station data as benchmark, predicted vs observed scatter plot can be created, along with a linear model and the $$R^2$$ value. I compare first the BREB against the EC data from the Ecosystem station:

```{r}
# line
model = lm(H_Wm2_BREB~H_Wm2, data = BREB_data)

# plot 
plot(H_Wm2_BREB~H_Wm2, data = BREB_data,
     ylab = "BREB sensible heat [Wm2]", 
     xlab = "EC sensible heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "darkred", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

Now, do the same for comparison of the ICOs Ecosystem station fluxes from 30m for the sensible heat flux, and my own calculated sensible heat flux from the sonic on 30m.

```{r}
# line
model = lm(H_Wm2_sonic_30m~H_Wm2, data = BREB_data)

# plot 
plot(H_Wm2_sonic_30m~H_Wm2, data = BREB_data, 
     ylab = "sonic sensible heat [Wm2]", 
     xlab = "EC sensible heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "green", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

Comparison of the latent heat flux from the BREb approach, and the latent heat flux from the ICOS Ecosystem station data:

```{r}
# linear model
model = lm(LE_Wm2_BREB~LE_Wm2, data = BREB_data)

# plot 
plot(LE_Wm2_BREB~LE_Wm2, data = BREB_data, 
     ylab = "BREB latent heat [Wm2]", 
     xlab = "EC latent heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "blue", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

# K-theory / Profile method

This is some code (wont run, since variable names changed) to calculate the eddy diffusivity for sensible heat $$K_H$$, which can be used later for the calculation of fluxes from simple gradients, without the need to use MOST Theory. Note, that the 118 in the calculation results from 148m-30m.

```{r}
radiation_filtered = radiation_filtered%>%
  mutate(
    K_H = - (sensible_heat_Wm2_Eddy_Cov_uncorrected)/(delta_T_C/118)
  )

# when now using K_H for calculating the sensible heat flux using K theory, this would be exactly the same result

# compare EC data from 30m height with ecosystem station measurements

```
