---
title: "Intercomparison of flux-estimation approaches"
author: "Felician Heim"
format: html
editor: visual
bibliography: references.bib
---

# Preparations and load data

First, packages are loaded.

```{r}
#| echo: false
library(tidyverse)
```

Second, the different data sets are loaded, which have been preprocessed (setting date and time, converting units and renaming of objects) in `preprocessing/...`:

-   EC data output from EddyPro (3 different heights, 30m, 70m, 148m) for the sensible heat flux and $K_H$, until now only year 2020 and 30m measurement height available

-   Profile measurements (slow) from 14 levels for temperature and humidity (and potentially other gases - Picarro data)

-   L2 ICOS Ecosystem station data (EC & radiation) for evaluation of fluxes (and provides ground pressure for obtaining water vapor pressure $e$)

```{r}
#| warning: false

#load EC data output first, here only 30m until now!!!
load("../data/processed/sonic_profile_data.RData")

# load the 14 level profile data for Ta and Humidity
load("../data/processed/slow_profile_data.RData")

# Ecosystem data 30m
load("../data/processed/Eco_data_30m.RData")

```

# Bowen ratio energy balance approach

The Bowen Ratio energy balance approach does not make any assumptions about turbulence, however, makes use of data on net radiation components as well as gradient measurements from at least two heights. The calculation is based on the Bowen Ration $Bo$, being defined as

$$
Bo = \gamma\frac{\Delta T}{\Delta e} 
$$

with $\Delta T$ and $\Delta e$ denoting the air temperature and water vapor difference between two heights and The psychrometric constant $$\gamma = 0.667 hPa/K$$ for air pressure $$p = 1013.25 hPa$$ and air temperature of 20ºC (other values are given in tables, see @micromet2026, page 120/121 ).

Then, the sensible and latent heat flux can be obtained from the following equation @foken2024, page 164:

$$
Q_H = (-Q_{Net}-Q_G)*\frac{Bo}{1+Bo}
$$

$$
Q_E = \frac{-Q_{Net}-Q_G}{1+Bo}
$$

here, $Q_{Net}$, $Q_H$, $Q_E$, and $Q_G$ denote the net radiation, sensible heat flux, latent heat flux and ground heat flux all in $W/m^2$, respectively. This means, the following data has to be available to calculate the the sensible and latent heat flux (in brackets where the data is coming from):

-   net radiation (Wm2) - from the `Eco_data_30m` dataset, so the ICOS Ecosystem station

-   ground heat flux (Wm2) - from the `Eco_data_30m` dataset, so the ICOS Ecosystem station

-   The $\Delta T$ and \$\\Delta e\$ from two measurement heights - from the `slow_profile_data` data set, using first 148m - 30m

    -   in order to obtain $e$ in $hPa$, one needs to convert the H2O mixing ratio in $mmol/mol$ first in $kg/kg$ by using $m[kg/kg] = m[mmol/mol]/(1000*0.622)$ (source/citation missing)

    -   and then, using the following equation @foken2024, page 41, Supplement 2.2, one can obtain the water vapor pressure $$e [hPa]= (m[kg/kg]*p[hPa])/(0.622)$$

    -   The gradient can then be determined as $\Delta e = e_{148m} - e_{30m}$

    -   This means, I need the ground pressure data from the Ecosystem Station in `Eco_data_30m` to calculate $e$

-   the gradient of the specific humidity $\Delta q$ is needed for the filter criterion proposed by @ohmura1982 for the BREB fluxes

First, all the needed data column are merged in one data frame:

```{r}
# from the sonic data, select only H value
sonic_profile_data = sonic_profile_data%>%
  filter(height == "30m")%>%
  rename(H_Wm2_sonic_30m = H_ensemble_mean)%>%
# keep only distinct rows (datetim is ther 4 times)
  distinct(datetime, .keep_all = TRUE)


slow_profile_data = slow_profile_data%>% select(datetime, 
                           Ta_dgC_148m, 
                            Ta_dgC_30m,
                            H2O_mmol_mol_148m, 
                            H2O_mmol_mol_30m)

Eco_data_30m = Eco_data_30m%>% select(datetime, 
                       H_Wm2, 
                       G_Wm2, 
                       LE_Wm2, 
                       R_Net_Wm2, 
                       P_ground_hPa)

# now merge, based on datetime
# use Eco_data as df to join in, since it contains complete 30min interval of 2020
BREB_data = Eco_data_30m%>%
  left_join(y = slow_profile_data, by = "datetime")%>%
  left_join(y = sonic_profile_data, by = "datetime")
```

Now, I can use the data to calculate the gradient $\Delta T$ and $\Delta e$, starting with the latter by make the unit conversion for mixing ratio in $mmol/mol$ to $kg/kg$, then $e[hPa]$.

```{r}
# calculate delta e
BREB_data = BREB_data%>%
  mutate(
    
    # convert from mmol/mol to kg/kg
    H2O_kg_kg_30m = H2O_mmol_mol_30m / 1000 * 0.622,
    H2O_kg_kg_148m = H2O_mmol_mol_148m / 1000 * 0.622,
    
    # calculate e in hPa
    e_148m_hPa = (P_ground_hPa * H2O_kg_kg_148m) / 0.622, 
    e_30m_hPa  = (P_ground_hPa * H2O_kg_kg_30m) / 0.622, 
    delta_e_hPa = e_148m_hPa - e_30m_hPa, 
    
    # calculate delta T
    delta_Ta_dgC = Ta_dgC_148m - Ta_dgC_30m
  )%>%
  # remove kg/kg columns
  select(-c(H2O_kg_kg_30m, 
            H2O_kg_kg_148m))

```

After this step, the formula for the Bowen ratio for the sensible and the latent heat can be used:

```{r}
# do the calculation for the latent and sensible heat flux
BREB_data = BREB_data%>%
  mutate(
    Bowen_ratio = 0.667 * (delta_Ta_dgC/delta_e_hPa),
    H_Wm2_BREB = (R_Net_Wm2-G_Wm2)*((Bowen_ratio)/(1+Bowen_ratio)),
    LE_Wm2_BREB = (R_Net_Wm2 - G_Wm2)/(1+Bowen_ratio)
  ) 
```

However, following @foken2024 on page 165, data must be filtered for quality reasons. Furthermore, The criteria proposed by @ohmura1982 is used for further filtering. Since the specific humidity $q$is needed for the criterion, $q$ is calculated using the formula from @foken2024, supplement 2.2 on page 54, with ground pressure from the Ecosystem station. Note, that I do not take into account the temperature dependency of the latent heat of vaporization lambda, but rather treat it as constant, equally the specific heat capacity at constant pressure $c_p$. Check the direction of -R_NET!

```{r}
# do quality filtering, based on Foken, Micrometeorology, page 65
BREB_data = BREB_data %>%
  mutate(
    H_Wm2_BREB  = if_else(Bowen_ratio > -1.25 
                                      & Bowen_ratio < -0.75, 
                                      true = NA, 
                                      false = H_Wm2_BREB),
    LE_Wm2_BREB = if_else(Bowen_ratio > -1.25 
                                     & Bowen_ratio < -0.75, 
                                   true = NA, 
                                   false = LE_Wm2_BREB)
  )

# calculate delta q in kg/kg
BREB_data= BREB_data%>%
  mutate(
    q_148m_kg_kg = 0.622*((e_148m_hPa)/(P_ground_hPa-0.378*e_148m_hPa)), 
    q_30m_kg_kg = 0.622*((e_30m_hPa)/(P_ground_hPa-0.378*e_30m_hPa)), 
    delta_q_kg_kg = q_148m_kg_kg - q_30m_kg_kg
  )

# filtering criteria based on Ohmura 1982 
lambda = 2.25e6 #in Joule per kg (although in reality dependend on T!)
c_p = 1005 #J/kg*K

BREB_data = BREB_data %>%
  mutate(
    valid =
      # one direction: both larger 0: >
      ((-R_Net_Wm2 - G_Wm2) > 0 
       & (lambda*delta_q_kg_kg + c_p*delta_Ta_dgC) > 0) | # or
      # now other direction <
      ((-R_Net_Wm2 - G_Wm2) < 0 
       & (lambda*delta_q_kg_kg + c_p*delta_Ta_dgC) < 0),

    # apply
    H_Wm2_BREB  = if_else(valid, true  = H_Wm2_BREB, false = NA),
    LE_Wm2_BREB = if_else(valid, true = LE_Wm2_BREB, false = NA)
  )

```

Now, the sensible and latent heat calculated using the BREB approach can be compared to the data from the Ecosystem station (EC system, processed using the ICOS protocol, 30m) and the sensible heat flux from the sonic from 30m processed by myself (EddyPro express mode for now).

Note, that this is only the data as it comes from Eddy pro (for the eddy data) - I did not do any filtering based on quality flags etc!

Using the Ecosystem station data as benchmark, predicted vs observed scatter plot can be created, along with a linear model and the $R^2$ value. I compare first the BREB against the EC data from the Ecosystem station:

```{r}
# line
model = lm(H_Wm2_BREB~H_Wm2, data = BREB_data)

# plot 
plot(H_Wm2_BREB~H_Wm2, data = BREB_data,
     ylab = "BREB sensible heat [Wm2]", 
     xlab = "EC sensible heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "darkred", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

Now, do the same for comparison of the ICOs Ecosystem station fluxes from 30m for the sensible heat flux, and my own calculated sensible heat flux from the sonic on 30m.

```{r}
# line
model = lm(H_Wm2_sonic_30m~H_Wm2, data = BREB_data)

# plot 
plot(H_Wm2_sonic_30m~H_Wm2, data = BREB_data, 
     ylab = "sonic sensible heat [Wm2]", 
     xlab = "EC sensible heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "green", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

Comparison of the latent heat flux from the BREb approach, and the latent heat flux from the ICOS Ecosystem station data:

```{r}
# linear model
model = lm(LE_Wm2_BREB~LE_Wm2, data = BREB_data)

# plot 
plot(LE_Wm2_BREB~LE_Wm2, data = BREB_data, 
     ylab = "BREB latent heat [Wm2]", 
     xlab = "EC latent heat [Wm2] ICOS Ecosystem station", 
     las = 1, pch = 1, cex = 0.5, xlim = c(-200, 830), ylim = c(-200, 830))

# add grid, abline and legend with R2
grid()
abline(0, 1, col = "black", lwd = 2, lty = 2)  
abline(model, col = "blue", lwd = 2)
legend("topleft", 
       legend = c(paste("R2 =", round(summary(model)$r.squared, 3))), bty = "n"
       )
```

# K-theory / Profile method

This is some code (wont run, since variable names changed) to calculate the eddy diffusivity for sensible heat $K_H$, which can be used later for the calculation of fluxes from simple gradients, without the need to use MOST Theory. Note, that the 118 in the calculation results from 148m-30m.

```{r}
radiation_filtered = radiation_filtered%>%
  mutate(
    K_H = - (sensible_heat_Wm2_Eddy_Cov_uncorrected)/(delta_T_C/118)
  )

# when now using K_H for calculating the sensible heat flux using K theory, this would be exactly the same result

# compare EC data from 30m height with ecosystem station measurements

```

# Footprint visualization

EddyPro gives as output the wind direction for the 30min (or other) averaging period, as well as the contribution of the different distances from the tower.

sonic_profile_data

```{r}
library(openair)

# need wind direction 
sonic_profile_data["wind_dir_[deg_from_north]"]

# and the contribution ### ATTENTION - might break because % is a dot
sonic_profile_data["x_10%_[m]"]
sonic_profile_data["x_30%_[m]"]
sonic_profile_data["x_50%_[m]"]
sonic_profile_data["x_70%_[m]"]
sonic_profile_data["x_90%_[m]"]

# and the wind speed
sonic_profile_data["wind_speed_[m+1s-1]"]


```

First, I can create a wind rose.

```{r}

wind_data = data.frame('wd' = as.numeric(sonic_profile_data[["wind_dir_[deg_from_north]"]])
, 
'ws' = as.numeric(sonic_profile_data[["wind_speed_[m+1s-1]"]]), 
date = date(sonic_profile_data[["datetime"]]))

# do for full data series or per month
polarFreq(wind_data)
polarFreq(wind_data, type = "month")


```

However, this is not a footprint of the measurements. This can be plotted from the multiplied wind direction and distance.

```{r}

wind_dir = data.frame(
  # need wind direction 
  dir = as.numeric(sonic_profile_data[["wind_dir_[deg_from_north]"]]), 

  # and the contribution for the respective percentages in distance meter
  perc_10 = as.numeric(sonic_profile_data[["x_10%_[m]"]]), 
  perc_30 = as.numeric(sonic_profile_data[["x_30%_[m]"]]), 
  perc_50 = as.numeric(sonic_profile_data[["x_50%_[m]"]]), 
  perc_70 = as.numeric(sonic_profile_data[["x_70%_[m]"]]), 
  perc_90 = as.numeric(sonic_profile_data[["x_90%_[m]"]]), 
  
  # date
  datetime = sonic_profile_data$datetime
)

wind_dir[wind_dir == -9999.0000] <- NA

# calculate footprint, convert from polar to cartesian coordinates
footprint <- wind_dir %>%
  mutate(
    wd_rad = dir * pi / 180,
    x = perc_90 * sin(wd_rad),
    y = perc_90 * cos(wd_rad)
  )

# -9999 --> NAs
footprint[footprint == -9999.0000] <- NA


# create a mean wind direction (circular mean)
mean_wind_dir <- function(wd) {
  wd_rad <- wd * pi / 180
  
  x <- mean(cos(wd_rad), na.rm = TRUE)
  y <- mean(sin(wd_rad), na.rm = TRUE)
  
  mean_rad <- atan2(y, x)
  mean_deg <- mean_rad * 180 / pi
  
  (mean_deg + 360) %% 360
}

# use function
wind_dir%>%
  group_by(week(datetime))%>%
  summarise(wind_circ_mean = mean_wind_dir(dir))%>%
  ungroup()

# plot
ggplot(footprint, aes(x = x, y = y)) +
  geom_point(alpha = 0.3) +
  coord_equal()+
  labs(color = "Percentile")


# 
library(sf)
# make coordinates from local coordinate system to global
tower_lonlat <- st_sfc(
  st_point(c(13.4189, 56.0976)),  # replace with your coordinates!
  crs = 4326
)

tower_utm <- st_transform(tower_lonlat, 32633) # UTM 33N

tower_coords <- st_coordinates(tower_utm)

tower_x <- tower_coords[1]
tower_y <- tower_coords[2]

# add the footprint in m
footprint_long <- footprint %>%
  mutate(
    x_map = x + tower_x,
    y_map = y + tower_y
  )

# delete NA
footprint_long = na.omit(footprint_long)

# convert to UTM 
footprint_sf <- st_as_sf(
  footprint_long,
  coords = c("x_map", "y_map"),
  crs = 32633
)

# make sf tower object
tower_sf <- st_sfc(
  st_point(c(tower_x, tower_y)),
  crs = 32633
)

ggplot() +
  geom_sf(data = footprint_sf, aes(color = "red", alpha = 0.5)) +
  geom_sf(data = tower_sf, color = "black", size = 3)
```

Make a background map:

```{r}

library(leaflet)

# transform to Wgs84
footprint_wgs84 <- st_transform(footprint_sf, 4326)
tower_wgs84     <- st_transform(tower_sf, 4326)


leaflet() %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addCircleMarkers(data = st_coordinates(footprint_wgs84), fillOpacity = 0.01, radius  = 0.1)%>%
  addCircleMarkers(
    data = tower_wgs84,
    color = "red",
    radius = 6,
    fillOpacity = 1
  )
```

# Comparison 30/70/148m latent heat flux

First, one can plot the ensembles for a specific time series for 30m and 70m. 148m height sis still to be added.

```{r}
#load data again
load("../data/processed/sonic_profile_data.RData")

# plot the different ensemble members and the mean
# select specific time frame

sonic_profile_data %>%
  filter(between(date(datetime),
                 ymd("2021-06-21"),
                 ymd("2021-06-26"))) %>%
  ggplot(aes(
    x = datetime,
    y = `H_[W+1m-2]`,
    color = height,
    group = interaction(height, folder)
  )) +
  geom_line() +
  theme_bw()

# compare the different height of sonic data
sonic_profile_data %>%
  group_by(datetime, height) %>%
  # take the mean of all the ensembles (grouped by height)
  summarise(`H_[W+1m-2]` = mean(`H_[W+1m-2]`, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = height,
    values_from = `H_[W+1m-2]`
  ) %>%
  ggplot(aes(x = `30m`, y = `70m`)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1.5) +
  geom_smooth(method = "lm", color = "red", size = 1) +
  ylab("H 70m [Wm2]") +
  xlab("H 30m [Wm2]") +
  theme_bw()


# make comparison of the detrending procedure: block and linear
sonic_profile_data%>%
  filter(between(date(datetime),
                 ymd("2021-01-21"),
                 ymd("2021-05-26")))%>%
  filter(height == "30m")%>%
  select(datetime, detrending, rotation, `H_[W+1m-2]`) %>%
  pivot_wider(
    names_from = detrending,
    values_from = `H_[W+1m-2]`
  )%>%
  ggplot()+
  geom_point(aes(x = block, y = linear))+
  ylab("H linear [Wm2]")+
  xlab("H block [Wm2]")+
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1.5)+
  geom_smooth(aes(x = block, y = linear), method = "lm", color = "red", size = 1)+
  facet_grid(~rotation)+
  theme_bw()

```

## look at u\*

Is there a relationship between turbulence strength and the flux strength (this should not be the case) for the eddy covariance fluxes?

Do this first for the whole data

```{r}
sonic_profile_data %>%
  distinct(datetime, .keep_all = TRUE)%>%
  filter(height == "30m") %>%
  ggplot(aes(x = `u*_[m+1s-1]`, y = H_ensemble_mean)) +
  geom_point(size = 0.5, alpha = 0.4) +
  geom_smooth(method = "loess", color = "red", size = 1) +
  ylab("H [Wm2]") +
  xlab("u* [m/s]") +
  
  theme_bw()
```

## optimizing the BREB approach

The BREB approach uses the gradient between different measurement heights (along with the eddy diffusivity $K$) to estimate the surface-atmosphere fluxes. However, having 14 measurement heights, the gradient is essentially a linear interpolation between two heights, while other values are neglected. To maximise the performance of the K-approach and also profile method, one can change the heights used to calculate the gradient. This can help to undersand the best (in terms of instrument setup) configuration of the method. Also, remaining differences between the BREB approach / profile method can be attributed solely to the method used, and not to possibly sub-optimal instrumental setup.

To do this, we use all (physically reasonable) combinations of measurement heights to calculate the gradient. This can be understood from a Machine Learning (ML) perspective as a type of hyperparameter tuning. The different levels used as a gradient are then compared based on the residuals (bias and spread).

The available measurement heights are the following:

```{r}

# define a function fro sensible and latent heat using BREB approach
BREB = function(H2O_mmol_mol_up, 
                          H2O_mmol_mol_down, 
                          Ta_dgC_up, 
                          Ta_dgC_down, 
                          P_ground_hPa, 
                          R_net_Wm2, 
                          G_Wm2, 
                          type = c("latent", "sensible")
                          ){
  
  # get the type of turbulent flux to be calculated
  type = match.arg(type)
  
  # convert from mmol/mol to kg/kg
  H2O_kg_kg_down = H2O_mmol_mol_down / 1000 * 0.622
  H2O_kg_kg_up = H2O_mmol_mol_up / 1000 * 0.622
    
  # calculate e in hPa
  e_hPa_up = (P_ground_hPa * H2O_kg_kg_up) / 0.622 
  e_hPa_down  = (P_ground_hPa * H2O_kg_kg_down) / 0.622
    
  # delta e
  delta_e_hPa = e_hPa_up - e_hPa_down 
    
  # calculate delta T
  delta_Ta_dgC = Ta_dgC_up - Ta_dgC_down
  
  ######### BREB sensible and latent heat calculation #########
  Bowen_ratio = 0.667 * (delta_Ta_dgC/delta_e_hPa)
  H_Wm2_BREB = (R_net_Wm2-G_Wm2)*((Bowen_ratio)/(1+Bowen_ratio))
  LE_Wm2_BREB = (R_net_Wm2 - G_Wm2)/(1+Bowen_ratio)
  
  ######### Quality filtering based on Foken, Micrometeorology, page 65 #########
  H_Wm2_BREB  = ifelse(Bowen_ratio > -1.25 
                        & Bowen_ratio < -0.75, 
                          yes = NA, 
                          no = H_Wm2_BREB)
  LE_Wm2_BREB = ifelse(Bowen_ratio > -1.25 
                          & Bowen_ratio < -0.75, 
                          yes = NA, 
                          no = LE_Wm2_BREB)
  
  ######### additional filtering criteria based on Ohmura 1982 #########
  # calculate delta q in kg/kg
  q_up_kg_kg = 0.622*((e_hPa_up)/(P_ground_hPa-0.378*e_hPa_up)) 
  q_down_kg_kg = 0.622*((e_hPa_down)/(P_ground_hPa-0.378*e_hPa_down))
  delta_q_kg_kg = q_up_kg_kg - q_down_kg_kg
  
  # set some variables
  lambda = 2.25e6 #in Joule per kg (although in reality dependend on T!)
  c_p = 1005 #J/kg*K
  
  valid =
      # one direction: both larger 0: >
      ((-R_net_Wm2 - G_Wm2) > 0 
       & (lambda*delta_q_kg_kg + c_p*delta_Ta_dgC) > 0) | # or
      # now other direction <
      ((-R_net_Wm2 - G_Wm2) < 0 
       & (lambda*delta_q_kg_kg + c_p*delta_Ta_dgC) < 0)

    # apply
    H_Wm2_BREB  = if_else(valid, true  = H_Wm2_BREB, false = NA)
    LE_Wm2_BREB = if_else(valid, true = LE_Wm2_BREB, false = NA)

  
  # return sensible heat as default
  if (type == "latent") {
    return(LE_Wm2_BREB)
  } else {
    return(H_Wm2_BREB)
  }
}

```

# tune the heights

```{r}

# heights
heights <- c(1,4,9,14,19,24,30,40,55,70,85,100,125,148)
pairs <- combn(heights, m = 2, simplify = FALSE) # get combos, result: c(down, up)
# create plot?
plot = FALSE

# storage for predictions
result = data.frame(datetime = BREB_data_tuning$datetime, 
                    LE_Wm2_truth = BREB_data_tuning$LE_Wm2_truth, 
                    H_Wm2_truth = BREB_data_tuning$H_Wm2_truth)
rmse_mean = c()
rmse_H = c()
rmse_LE = c()



for(i in 1:length(pairs)){
  
  h_down = pairs[[i]][1]
  h_up = pairs[[i]][2]
  
  
  # APPLY FUNCTION for H for 
  result[[paste0("H_", h_down, "_", h_up)]] = 
    BREB(
      H2O_mmol_mol_up = BREB_data_tuning[[paste0("H2O_", h_up, "m")]], 
      H2O_mmol_mol_down = BREB_data_tuning[[paste0("H2O_", h_down, "m")]], 
      Ta_dgC_up = BREB_data_tuning[[paste0("Ta_", h_up, "m")]], 
      Ta_dgC_down = BREB_data_tuning[[paste0("Ta_", h_down, "m")]], 
      P_ground_hPa = BREB_data_tuning$P_ground_hPa, 
      R_net_Wm2 = BREB_data_tuning$R_Net_Wm2, 
      G_Wm2 = BREB_data_tuning$G_Wm2, 
      type = "sensible"
    )
  
  # apply function for LE
  result[[paste0("LE_", h_down, "_" , h_up)]] = 
    BREB(
      H2O_mmol_mol_up = BREB_data_tuning[[paste0("H2O_", h_up, "m")]], 
      H2O_mmol_mol_down = BREB_data_tuning[[paste0("H2O_", h_down, "m")]], 
      Ta_dgC_up = BREB_data_tuning[[paste0("Ta_", h_up, "m")]], 
      Ta_dgC_down = BREB_data_tuning[[paste0("Ta_", h_down, "m")]], 
      P_ground_hPa = BREB_data_tuning$P_ground_hPa, 
      R_net_Wm2 = BREB_data_tuning$R_Net_Wm2, 
      G_Wm2 = BREB_data_tuning$G_Wm2, 
      type = "latent"
    )
  
  # plot the results and compare to EC Eco 30m 
  model_LE = lm(
    as.formula(paste0("LE_", h_down, "_", h_up, " ~ LE_Wm2_truth")),
    data = result)
  
  # get results & calculate RMSE LE
  y <- result[[paste0("LE_", h_down, "_", h_up)]]
  x <- result$LE_Wm2_truth
  rmse_LE <- append(rmse_LE, sqrt(mean((y - x)^2, na.rm = TRUE)))
  
  ###### for H #######
  model_H = lm(
    as.formula(paste0("H_", h_down, "_", h_up, " ~ H_Wm2_truth")),
    data = result)
  
  # get results and calculate RMSE for H
  y <- result[[paste0("H_", h_down, "_", h_up)]]
  x <- result$H_Wm2_truth
  rmse_H <- append(rmse_H, sqrt(mean((y - x)^2, na.rm = TRUE)))
  
  # store mean
  rmse_mean = append(rmse_mean, mean(c(rmse_H[i], rmse_LE[i])))
  
  
  if(plot == TRUE){
    par(mfrow = c(1, 2))
    
    # plot H
    plot(
      x = result$H_Wm2_truth,
      y = result[[paste0("H_", h_down, "_", h_up)]],
      ylab = paste0("LE modeled ", h_down, "m & ", h_up, "m [Wm2]"),
      xlab = "H EC eco 30m [Wm2]",
      main = paste("H", h_down, "-", h_up), 
      cex = 0.5,
      ylim = c(-250, 600)
    )
    
    # grid, legend abline etc
    grid()
    abline(0, 1, lwd = 2, lty = 2)
    abline(model_H, lwd = 2, col = "red")
    legend(
      "topleft",
      legend = paste("R2 =", round(summary(model_H)$r.squared, 3)),
      bty = "n")
    
    # plot LE
    plot(
      x = result$LE_Wm2_truth,
      y = result[[paste0("LE_", h_down, "_", h_up)]],
      ylab = paste0("LE modeled ", h_down, "m & ", h_up, "m [Wm2]"),
      xlab = "LE EC eco 30m [Wm2]",
      main = paste("LE", h_down, "-", h_up),
      cex = 0.5,
      ylim = c(-250, 600)
    )
    
    # grid, legend abline etc
    grid()
    abline(0, 1, lwd = 2, lty = 2)
    abline(model_LE, lwd = 2, col = "red")
    legend(
      "topleft",
      legend = paste("R2 =", round(summary(model_LE)$r.squared, 3)),
      bty = "n")
    
    par(mfrow = c(1, 1))
  }
  
}

# extract results

h_down_vec <- sapply(pairs, `[`, 1)
h_up_vec   <- sapply(pairs, `[`, 2)

res <- data.frame(
  h_down = h_down_vec,
  h_up   = h_up_vec,
  rmse_mean = rmse_mean, 
  rmse_H = rmse_H, 
  rmse_LE = rmse_LE
)

# get mininum rmse
res[which(res$rmse_mean == min(res$rmse_mean)),] # mean
res[which(res$rmse_H == min(res$rmse_H)),] # H
res[which(res$rmse_LE == min(res$rmse_LE)),] # LE


# 19 and 40m for overall min(rmse_H & rmse_LE)
# 14 and 40m for H
# 19 & 125 for LE



```

### stability parameter

Probably, the accuracy and precision of the modeled fluxes is influenced by the stability of the ABL. To quantify the stability, the stability parameter $z/\Lambda_*$ can be used and follwing @sorbjan2010 be classified into several classes, and then plotted within the observed vs predicted scatter plot for H and LE. However, it does not seem that

```{r}
# define a function to classify stability following Sorbjan et al 2010

classify_stability <- function(zeta) {
  ifelse(is.na(zeta), NA,
  ifelse(zeta < -1, "xu",
  ifelse(zeta < -0.6, "vu",
  ifelse(zeta < -0.2, "u",
  ifelse(zeta < -0.02, "wu",
  ifelse(zeta < 0.02, "n",
  ifelse(zeta < 0.2, "ws",
  ifelse(zeta < 0.6, "s",
  ifelse(zeta < 1, "vs",
         "xs")))))))))
}

# apply the classification
sonic_profile_data = sonic_profile_data%>%
  mutate(stability_classes = classify_stability(`(z-d)/L_[#]`))

# merge with results
result = left_join(x = result, 
          y = sonic_profile_data%>%select(datetime, stability_classes), 
          by = "datetime")

# plot for H and LE
ggplot(result)+
  geom_point(aes(x = H_Wm2_truth, y = H_14_40, color = stability_classes), size = .4, alpha = 0.3)+
  ylab("H 19m & 40m [Wm2]")+
  xlab("H EC eco [Wm2]")+
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1.5)+
  geom_smooth(aes(x = H_Wm2_truth, y = H_14_40, color = stability_classes), method = "lm", size = 1)+
  theme_bw()

ggplot(result)+
  geom_point(aes(x = LE_Wm2_truth, y = LE_14_40), size = .4, alpha = 0.3)+
  ylab("LE 19m & 40m [Wm2]")+
  xlab("LE EC eco [Wm2]")+
  geom_abline(intercept = 0, slope = 1, color = "black", size = 1.5)+
  geom_smooth(aes(x = LE_Wm2_truth, y = LE_14_40, color = stability_classes), method = "lm", size = 1)+
  theme_bw()


```

Inspect whether the stability classes make sense:

-   classify the profiles in the different classes, then plot them.

Yes, I think roughly, they do make sense. However, it does not classify / predict the classes very well, and there are always lots of different shapes of profiles in there. Note also, that the best prediction (based on RMSE) has been achieved using the measurement heights of 14/19m and 40m, indicating that the 10m above the canopy are most important for estimating the fluxes

```{r}
slow_profile_data%>%
  left_join(sonic_profile_data%>%select(stability_classes, datetime), 
           by = "datetime")%>%
  select(-c(date.y, time.y, doy.y))%>%
  filter(stability_classes == "xs")%>%
  pivot_longer(
    cols = matches("^Ta_"),
    names_to = "height_m",
    values_to = "value"
  ) %>%
  mutate(
    # make numeric from character
    height_m = str_remove(height_m, "Ta_"), 
    height_m = str_remove(height_m, "m"), 
    height_m = as.numeric(height_m))%>%
  filter(between(date(datetime),
                 ymd("2021-06-01"),
                 ymd("2021-07-01"))) %>%
  filter(height_m <50)%>%
  arrange(height_m)%>%
  ggplot(aes(
    x = value,
    y = height_m,
    group = datetime
  )) +
  geom_path() +
  ylab("Height [m]") +
  xlab("Air temperature [°C]") +
  theme_bw()
  


```
